---
title: "A Walkthrough of blueprintr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Walkthrough of blueprintr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
div.vg-warning {
  background-color:#FFFCE8;
  padding:5px;
}

div.vg-warning span:before {
  content: "⚠️ Warning: ";
}

div.vg-info {
  background-color:#EBFBFF;
  padding:5px;
}

div.vg-info span:before {
  content: "💡 Note: ";
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

set.seed(0x7bae)

cache_location <- tempdir()

drake::clean(cache = drake::drake_cache(cache_location))
```

blueprintr is a companion to [drake](https://github.com/ropensci/drake) that focuses on documenting and testing tabular data. Whereas drake manages the workflow execution, blueprintr defines a collection of steps that need to be run in a drake workflow.

# Basic Use

The first, and recommended, step is to attach blueprintr to your R session with `library()`.

```{r setup}
library(blueprintr)
```

In a [drake project](https://books.ropensci.org/drake/projects.html), all packages that you want attached are declared in a `"packages.R"` file. This `library(blueprintr)` command should go there.

blueprintr is built around "blueprints." Our first blueprint will be a blueprint for `mtcars`:

```{r}
blueprint(
  "mtcars_dat",
  description = "The famous mtcars dataset",
  command = {
    mtcars
  }
)
```

All blueprints have

* A name (the first argument) for the _target_ dataset.
* A description or brief summary of what the target is. Can be `NULL`.
* A command, which is a quoted statement that has the code for building this target.
* A metadata location, which is a path to where the target metadata is saved.

<div class="vg-warning"><span>the blueprint name is "mtcars_dat" rather than "mtcars". If the two had the same name, drake would determine that the blueprint has a _circular dependency_ (it depends on itself). To avoid this, blueprints should not have the same names as global variables, like `mtcars`.</span></div>

To get this loaded into a drake plan, we need to _attach_ it to an already existing plan using

```r
attach_blueprint(plan, blueprint)
attach_blueprints(plan, ...)
```

<div class="vg-info"><span>`attach_blueprints` accepts "[tidy dots](https://adv-r.hadley.nz/quasiquotation.html#tidy-dots)", so if you have a `list()` of blueprints, you can "splat" all of those blueprints into `attach_blueprints` like `attach_blueprints(plan, !!!list_of_blueprints)`</span></div>

If you don't have an existing plan, you can create one with

```r
plan_from_blueprint(blueprint)
```

For now, we'll use an already existing plan, which is probably what you'll have most of the time.

```{r, include=FALSE}
existing_plan <- drake::drake_plan(initial_vector = runif(1000), squared = initial_vector ^ 2)
```

```{r}
attach_blueprint(
  existing_plan,
  blueprint(
    "mtcars_dat",
    description = "The famous mtcars dataset",
    command = {
      mtcars
    }
  )
)
```

blueprintr creates five "steps" (targets) per blueprint:

Target name             | Description
------------------------|--------------
`{blueprint}_initial`   | The result of running the blueprint's `command`
`{blueprint}_blueprint` | A copy of the blueprint to be used throughout the plan
`{blueprint}_meta`      | A copy of the dataset metadata --- if the metadata file doesn't exist, it will be created in this step
`{blueprint}_checks`    | Runs all checks on the `{blueprint}_initial` target
`{blueprint}`           | The built dataset after running some cleanup tasks

At this point, you're able to run `drake::make()` on this plan!

<div class="vg-warning"><span>when writing other targets in your plan, it is advised to **not** refer to the `{blueprint}_initial` step since it could have problems which are discovered in the `{blueprint}_checks` step.</span></div>

```{r delete_cache, include=FALSE}
unlink(cache_location, recursive = TRUE)
```

