---
title: "A Walkthrough of blueprintr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blueprintr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

blueprintr is a companion to [drake](https://github.com/ropensci/drake) that focuses on documenting and testing tabular data. Whereas drake manages the workflow execution, blueprintr defines a collection of steps that need to be run in a drake workflow.

# Basic Use

The first, and recommended, step is to attach blueprintr to your R session with `library()`.

```{r setup}
library(blueprintr)
```

In a [drake project](https://books.ropensci.org/drake/projects.html), all packages that you want attached are declared in a `"packages.R"` file. This `library(blueprintr)` command should go there.

blueprintr is built around "blueprints." Our first blueprint will be a blueprint for `mtcars`:

```{r}
blueprint(
  "mtcars",
  description = "The famous mtcars dataset",
  command = {
    mtcars
  }
)
```

All blueprints have

* A name (the first argument) for the _target_ dataset.
* A description or brief summary of what the target is. Can be `NULL`.
* A command, which is a quoted statement that has the code for building this target.
* A metadata location, which is a path to where the target metadata is saved.

